\documentclass[letterpaper,conference,fleqn]{IEEEtran}
%\documentclass[11pt,letterpaper,conference,fleqn]{IEEEtran}

\usepackage{times,epsfig,graphicx,cite,subfig,mathtools,amssymb}
\usepackage{url}

\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

% \pagestyle{empty}
%       \setlength{\parskip}{12pt}
%       \renewcommand{\baselinestretch}{1.1}

\begin{document}

% ------------------------------------------------------------
%\newcommand {\softNewpage}     {\newpage}
\newcommand {\softNewpage}      {}

%\newcommand* {\OmitProof} {}            % Expose this line to omit the
     % \ifdefined\OmitProof		% part after else
     % ...
     % \else
     % ...
     % \fi
% ------------------------------------------------------------
\newtheorem{lemma}{Lemma}
% \newtheorem{thm}{Theorem}
\newtheorem{thm}{{\break \noindent {\bf Theorem}}}

\newcommand{\loc}   	{ {\mathrm {Loc}} }
\newcommand{\ACONN}   { {\mathrm {A\mbox{-}CONN}} }
\newcommand{\SCONN}   { {\mathrm {S\mbox{-}CONN}} }

\newcommand{\fConn}   	{ {\mathrm {Conn}} }

\newcommand{\nReq}    { {n_{req}} }
\newcommand{\boldS}   { \mathbf{S}}
\newcommand{\DCH}     { {DCH} }

%% -- older paper --

\newcommand{\starEqual}   { {\;{\scriptstyle *} \! =} }
\newcommand{\plusEqual}   { {\;{\scriptstyle +} \! =} }
%\newcommand{\plusplus}    { {\scriptstyle \!+\!+} }

%\newcommand{\starEqual}   { {\; {\mathtt *=}} }
%\newcommand{\plusEqual}   { {\; \mathtt{+=}} }

% --------------------
\newcommand{\set}[1]    {{ \{ #1 \} }}
\newcommand{\iin}[1]    {\hspace*{#1in}}

\newcommand{\ol}[1]     {\overline{#1}}

\newcommand{\Prob}[1]   { {\bf \mathrm{Prob}} \left[ #1 \right] }
\newcommand{\aPr}	{ {\mathrm  {Pr}} }

\newcommand {\nwline} {\hfill\break}
% ------------------------------------------------------------
%	\pagestyle{myheadings}
%	\markboth{{\tt {E. Elmallah (draft)} \hfill}}
%	          {{\tt {E. Elmallah (draft)} \hfill}}
%
%	\begin{minipage}[t]{4in} \vspace*{-.5in}
%	{\small \tt submitted to ....} \vspace{.5in}
%	\end{minipage}
%
% ------------------------------------------------------------
\title {
       Tree Bound on Probabilistic Connectivity of Underwater Sensor Networks
}
\author {
        \IEEEauthorblockN{Md Asadul Islam}
	\IEEEauthorblockA{
                Department of Computing Science \\
                University of Alberta \\
                Edmonton, T6G 2E8, Canada\\
                E-Mail: mdasadul@ualberta.ca
        }
	\and
        \IEEEauthorblockN{Ehab S. Elmallah}
	\IEEEauthorblockA{
                Department of Computing Science \\
                University of Alberta \\
                Edmonton, T6G 2E8, Canada\\
                E-Mail: elmallah@ualberta.ca
        }
}
%	\date{March 21, 2014}
%	\thanks{This research is supported by NSERC Canada under grant
%		number OGP 36899.}
\maketitle
\thispagestyle{empty}
% --------------------
%
\begin{abstract}
This paper considers Underwater Sensor Networks (UWSNs) with unanchored
nodes that can move freely with water currents.
%
Thus, node locations at any instant can only be specified probabilistically.
%
When connectivity among some of the sensor nodes is required to perform
a given function, the problem of estimating the likelihood that the network
achieves such connectivity arises.
%
Our work here formulates a parameterized probabilistic connectivity
problem that serves this purpose when the network contains both sensor nodes
and relay nodes.
%
We show an exact dynamic programming algorithm for solving the problem
on networks with tree-like structure; 
the algorithm yields lower bounds on the solution of
the problem on any arbitrary network.
%
The obtained simulation results investigate the gap between our obtained
lower bounds and exact solutions on small networks, as well as
the usefulness of our method in analyzing the effect of adding relay nodes
to the network.
\end{abstract}
%
% ------------------------------
\ifdefined\OmitProof
%
\else
\nwline
\begin{IEEEkeywords}
Underwater sensor networks,
probabilistic connectivity,
probabilistic graphs.
\end{IEEEkeywords}
\fi
% ------------------------------
% ------------------------------------------------------------

%\section{Introduction}  {\bf To be completed}
\input{sec-intro.tex}

%\newpage
\input{sec-model.tex}

%\section{Main Algorithm} {\bf To be completed}
\input{sec-alg}

%\section{Simulation Results} {\bf To be completed}
\input{sec-sim.tex}

% ------------------------------------------------------------
\section {Concluding Remarks}
%
Quantifying the likelihood that a sufficient number of sensor nodes
is connected to a sink node in a given UWSN is a challenging problem for
networks with semi-mobile and mobile nodes.
%
Our work here adopts a flexible probabilistic graph model to formulate
a class of parameterized network connectivity problems.
%
The problem setting allows the network to utilize both sensor nodes
and relay nodes.
%
For scenarios where the model is exact, we show that the exact
probabilistic connectivity can be computed efficiently for tree-like
networks.
%
Our simulation results show the usefulness of the algorithm in
computing efficient lower bounds on the solution of any arbitrary
network. Scenarios with deployed relay nodes are also analyzed.
%
Future work will consider obtaining stronger bounds of the formulated
problems.

% ------------------------------------------------------------

%\nwline
%\centerline{\sc Acknowledgment}
%\section*{Acknowledgment}
% This research is supported by NSERC Canada.
% ------------------------------------------------------------
%	\newpage

% Styles: latex8, unsrt, IEEEtranS.bst
%
\bibliographystyle{IEEEtran}
\bibliography{tree-UWSN}

%% \bibliography{exposure,00bib/mypapers,00bib/wireless-sensor,00bib/rel}

%% \bibliography{00bib/mypapers,00bib/qos,00bib/rel,00bib/general,00bib/wireless-sensor}

% ============================================================
\end{document}
